Weather covariance matrices as U_hyp for GR50

### 2021-01-26 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Finish mash Results section writeup

```{r functions to plot svd of data driven covar matrices}
get_svd_df <- function(m, i = NULL, eigen_v = 1){
  Ulist <- switchgrassGWAS:::get_Ulist(m)
  stopifnot(eigen_v %in% c(1,2,3,4,5))
  if(is.null(i)){
    i_v <- which(names(Ulist) %in% names(Ulist)[get_U_by_mass(m, thresh = 0.1)])
    i = i_v[1]
  } else if(is.character(i)){
    i <- which(names(Ulist) %in% i)
  }
  stopifnot(i %in% 1:length(Ulist))

  b1 <- svd(Ulist[[i]])
  PVE_EV <- b1$d[eigen_v] / sum(b1$d) # Proportion of variance explained by this eigenvector
  ggb1 <- tibble(`Eigenvector` = b1$v[,eigen_v]/Ulist[[i]][which.max(abs(Ulist[[i]][,eigen_v])),eigen_v],
       Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_and_Midwest_")) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatMidwest_")) %>%
  separate(Phenotype, into = c("Phenotype", "Garden"), sep = -3) %>%
  mutate(Phenotype = str_remove(Phenotype, "_$"),
         Garden = str_remove(Garden, "^_"),
         PVE_EV = PVE_EV)
  ggb1$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
  return(ggb1)
}

mash_plot_svd <- function(m, i = NULL, eigen_v = 1){
  b1_df <- get_svd_df(m = m, i = i, eigen_v = eigen_v)
  if(eigen_v == 1){
    ylab <- paste0("First eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  } else if(eigen_v == 2){
    ylab <- paste0("Second eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  } else if(eigen_v == 3){
    ylab <- paste0("Third eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  } else if(eigen_v == 4){
    ylab <- paste0("Fourth eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  } else if(eigen_v == 5){
    ylab <- paste0("Fifth eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  }
  ggb1 <- b1_df %>%
    ggplot(aes(x = Garden, y = `Eigenvector`)) +
    switchgrassGWAS::theme_oeco + 
    geom_bar(stat = "identity", position = "dodge") +
    theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'),
          axis.text.x = element_text(hjust = 1, angle = 45)) +
    geom_hline(yintercept = 0, linetype = 2) +
    scale_color_viridis_d(option = "B", end = 0.7, 
                        direction = -1) +
    scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, 
                         direction = -1)  +
    labs(y = ylab)
  return(ggb1)
}
```

@@

@@

f3c <- mash_plot_pairwise_sharing(corrmatrix = gxe_gg$S_AP, reorder = FALSE)
f3d <- mash_plot_pairwise_sharing(corrmatrix = gxe_gg$S_DS, reorder = FALSE)
f3g <- mash_plot_pairwise_sharing(corrmatrix = gxe_mg$S_AP, reorder = FALSE)
f3h <- mash_plot_pairwise_sharing(corrmatrix = gxe_mg$S_DS, reorder = FALSE)
f3k <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamg$S_AP, reorder = FALSE)
f3l <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamg$S_DS, reorder = FALSE)
f3o <- mash_plot_pairwise_sharing(corrmatrix = gxe_gf$S_AP, reorder = FALSE)
f3p <- mash_plot_pairwise_sharing(corrmatrix = gxe_gf$S_DS, reorder = FALSE)
f3s <- mash_plot_pairwise_sharing(corrmatrix = gxe_mf$S_AP, reorder = FALSE)
f3t <- mash_plot_pairwise_sharing(corrmatrix = gxe_mf$S_DS, reorder = FALSE)
f3w <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamf$S_AP, reorder = FALSE)
f3x <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamf$S_DS, reorder = FALSE)
```
put labels in latitudinal order
get rid of all the STand_Bhat_pop_GR50_ garbage

```{r}
ggdraw() +
  draw_plot(f3a, x = 0, y = 0.825, width = 0.33, height = 0.175) +
  draw_plot(f3b, x = 0, y = 0.65, width = 0.33, height = 0.175) +
  draw_plot(f3d$gg_corr, x = 0.165, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3c$gg_corr, x = 0, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3e, x = 0.33, y = 0.825, width = 0.33, height = 0.175) +
  draw_plot(f3f, x = 0.33, y = 0.65, width = 0.33, height = 0.175) +
  draw_plot(f3h$gg_corr, x = 0.165+.33, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3g$gg_corr, x = 0+.33, y = 0.5, width = 0.165, height = 0.15, scale = 2)  +
  draw_plot(f3i, x = 0.66, y = 0.825, width = 0.33, height = 0.175) +
  draw_plot(f3j, x = 0.66, y = 0.65, width = 0.33, height = 0.175) +
  draw_plot(f3l$gg_corr, x = 0.165+.66, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3k$gg_corr, x = 0+.66, y = 0.5, width = 0.165, height = 0.15, scale = 2) +  ## greenup above here, flowering below
  draw_plot(f3m, x = 0, y = 0.825-0.5, width = 0.33, height = 0.175) +
  draw_plot(f3n, x = 0, y = 0.65-0.5, width = 0.33, height = 0.175) +
  draw_plot(f3p$gg_corr, x = 0.165, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3o$gg_corr, x = 0, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3q, x = 0.33, y = 0.825-0.5, width = 0.33, height = 0.175) +
  draw_plot(f3r, x = 0.33, y = 0.65-0.5, width = 0.33, height = 0.175) +
  draw_plot(f3t$gg_corr, x = 0.165+.33, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3s$gg_corr, x = 0+.33, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2)  +
  draw_plot(f3u, x = 0.66, y = 0.825-0.5, width = 0.33, height = 0.175) +
  draw_plot(f3v, x = 0.66, y = 0.65-0.5, width = 0.33, height = 0.175) +
  draw_plot(f3x$gg_corr, x = 0.165+.66, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3w$gg_corr, x = 0+.66, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2)

save_plot(filename = "Figure_3_Data_driven_Covariance_matrices.png", plot = last_plot(), base_height = 8, base_width = 9)
```

>>>>

```{r functions to plot svd of data driven covar matrices}
m <- m_midw_fl_hyp
get_svd_df <- function(m, i = NULL, eigen_v = 1){
  Ulist <- switchgrassGWAS:::get_Ulist(m)
  stopifnot(eigen_v %in% c(1,2,3,4,5))
  if(is.null(i)){
    i_v <- which(names(Ulist) %in% names(Ulist)[get_U_by_mass(m, thresh = 0.1)])
    i = i_v[1]
  } else if(is.character(i)){
    i <- which(names(Ulist) %in% i)
  }
  stopifnot(i %in% 1:length(Ulist))

  b1 <- svd(Ulist[[i]])
  PVE_EV <- b1$d[eigen_v] / sum(b1$d) # Proportion of variance explained by this eigenvector
  ggb1 <- tibble(`Eigenvector` = b1$v[,eigen_v]/Ulist[[i]][which.max(abs(Ulist[[i]][,eigen_v])),eigen_v],
       Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_and_Midwest_")) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatMidwest_")) %>%
  separate(Phenotype, into = c("Phenotype", "Garden"), sep = -3) %>%
  mutate(Phenotype = str_remove(Phenotype, "_$"),
         Garden = str_remove(Garden, "^_"),
         PVE_EV = PVE_EV,
         Garden = case_when(grepl("^0_M$", Garden) ~ "MI",
                            TRUE ~ Garden))
  ggb1$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
  return(ggb1)
}

mash_plot_svd <- function(m, i = NULL, eigen_v = 1){
  b1_df <- get_svd_df(m = m, i = i, eigen_v = eigen_v)
  if(eigen_v == 1){
    ylab <- paste0("First eigenvector\n(PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  } else if(eigen_v == 2){
    ylab <- paste0("Second eigenvector\n(PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  } else if(eigen_v == 3){
    ylab <- paste0("Third eigenvector\n(PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  } else if(eigen_v == 4){
    ylab <- paste0("Fourth eigenvector\n(PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  } else if(eigen_v == 5){
    ylab <- paste0("Fifth eigenvector\n(PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  }
  ggb1 <- b1_df %>%
    ggplot(aes(x = Garden, y = `Eigenvector`)) +
    switchgrassGWAS::theme_oeco + 
    geom_bar(stat = "identity", position = "dodge") +
    theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'),
          axis.text.x = element_text(hjust = 1, angle = 45)) +
    geom_hline(yintercept = 0, linetype = 2) +
    scale_color_viridis_d(option = "B", end = 0.7, 
                        direction = -1) +
    scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, 
                         direction = -1)  +
    labs(y = ylab)
  return(ggb1)
}
```

@@

```{r}
corrmatrix <-  gxe_mf$S_AP
pvdiv_plot_cor <- function(corrmatrix, upper = NULL, legend = FALSE){
  U1 <- corrmatrix

  U1 <- as_tibble(U1, rownames = "rowU", .name_repair = "unique") %>%
          pivot_longer(cols = -.data$rowU, names_to = "colU",
                       values_to = "covar") %>%
          filter(!is.na(.data$covar)) %>%
    mutate(rowU = str_remove(rowU, 
                             "Stand_BhatGulf_and_Midwest_")) %>%
    mutate(rowU = str_remove(rowU, "Stand_BhatGulf_")) %>%
    mutate(rowU = str_remove(rowU, "Stand_BhatMidwest_")) %>%
    mutate(rowU = str_remove(rowU, "GR50_")) %>%
    mutate(rowU = str_remove(rowU, "FL50_")) %>%
    mutate(colU = str_remove(colU, 
                             "Stand_BhatGulf_and_Midwest_")) %>%
    mutate(colU = str_remove(colU, "Stand_BhatGulf_")) %>%
    mutate(colU = str_remove(colU, "Stand_BhatMidwest_")) %>%
    mutate(colU = str_remove(colU, "GR50_")) %>%
    mutate(colU = str_remove(colU, "FL50_")) %>%
    mutate(rowU = case_when(grepl("^M$", rowU) ~ "MI",
                            TRUE ~ rowU),
           colU = case_when(grepl("^M$", colU) ~ "MI",
                            TRUE ~ colU),)
  U1$colU <- factor(U1$colU, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
  U1$rowU <- factor(U1$rowU, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
  U1 <- arrange(U1, rowU, colU) %>%
    pivot_wider(names_from = colU, values_from = covar) 
  for(i in 1:nrow(U1)){
    for(j in 2:ncol(U1)){
      if(i <= j-1){
        U1[i, j] <- NA
      }
    }
  }
  U1 <- U1 %>%
    pivot_longer(cols = -.data$rowU, names_to = "colU",
                       values_to = "covar")
  U1$colU <- factor(U1$colU, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD")) 
  if(is.null(upper)){
  upper <- max(U1$covar, na.rm = TRUE)+5
  }
  U1_covar <- U1 %>%
    filter(!is.na(.data$covar)) %>%
    ggplot(aes(x = .data$rowU, y = .data$colU)) +
    switchgrassGWAS::theme_oeco +
    geom_tile(aes(fill = .data$covar)) +
    scale_fill_viridis_c(option = "B", limits = c(0, upper)) +
    #geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    xlab("") + ylab("") 
  if(legend == TRUE){
    U1_covar <- U1_covar +
      theme(legend.position = c(0.2, 0.9)) +
      geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey")
  }
  return(U1_covar)
}

@@

f3c <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_AP, upper = 5500, legend = TRUE)
f3c_leg <- get_legend(f3c)
f3c <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_AP, upper = 5500, legend = FALSE)
f3d <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_DS, upper = 5500, legend = FALSE)

f3g <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_AP, upper = 7777, legend = TRUE)
f3g_leg <- get_legend(f3g)
f3g <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_AP, upper = 7777, legend = FALSE)
f3h <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_DS, upper = 7777, legend = FALSE)

f3k <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_AP, upper = 7580, legend = TRUE)
f3k_leg <- get_legend(f3k)
f3k <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_AP, upper = 7580, legend = FALSE)
f3l <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_DS, upper = 7580,  legend = FALSE)

f3o <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_AP, upper = 1850, legend = TRUE)
f3o_leg <- get_legend(f3o)
f3o <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_AP, upper = 1850, legend = FALSE)
f3p <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_DS, upper = 1850, legend = FALSE)

f3s <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_AP, upper = 3636, legend = TRUE)
f3s_leg <- get_legend(f3s)
f3s <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_AP, upper = 3636, legend = FALSE)
f3t <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_DS, upper = 3636, legend = FALSE)

f3w <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_AP, upper = 13438, legend = TRUE)
f3w_leg <- get_legend(f3w)
f3w <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_AP, upper = 13438, legend = FALSE)
f3x <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_DS, upper = 13438, legend = FALSE)
```
put labels in latitudinal order
get rid of all the STand_Bhat_pop_GR50_ garbage

```{r}
ggdraw() +
  draw_plot(f3a, x = 0, y = 0.84, width = 0.33, height = 0.16) +
  draw_plot(f3b, x = 0, y = 0.68, width = 0.33, height = 0.16) +
  draw_plot(f3d, x = 0.165, y = 0.5, width = 0.165, height = 0.18) +
  draw_plot(f3c, x = 0, y = 0.5, width = 0.165, height = 0.18) +
  draw_plot(f3c_leg, x = 0.095, y = 0.595, width = 0.1, height = 0.1) +
  draw_plot(f3e, x = 0.33, y = 0.84, width = 0.33, height = 0.16) +
  draw_plot(f3f, x = 0.33, y = 0.68, width = 0.33, height = 0.16) +
  draw_plot(f3h, x = 0.165+.33, y = 0.5, width = 0.165, height = 0.18) +
  draw_plot(f3g, x = 0+.33, y = 0.5, width = 0.165, height = 0.18)  +
  draw_plot(f3g_leg, x = 0.095+.33, y = 0.595, width = 0.1, height = 0.1) +
  draw_plot(f3i, x = 0.66, y = 0.84, width = 0.33, height = 0.16) +
  draw_plot(f3j, x = 0.66, y = 0.68, width = 0.33, height = 0.16) +
  draw_plot(f3l, x = 0.165+.66, y = 0.5, width = 0.165, height = 0.18) +
  draw_plot(f3k, x = 0+.66, y = 0.5, width = 0.165, height = 0.18) +  
  draw_plot(f3k_leg, x = 0.095+.66, y = 0.595, width = 0.1, height = 0.1) +## greenup above here, flowering below
  draw_plot(f3m, x = 0, y = 0.84-0.5, width = 0.33, height = 0.16) +
  draw_plot(f3n, x = 0, y = 0.68-0.5, width = 0.33, height = 0.16) +
  draw_plot(f3p, x = 0.165, y = 0.5-0.5, width = 0.165, height = 0.18) +
  draw_plot(f3o, x = 0, y = 0.5-0.5, width = 0.165, height = 0.18) +  
  draw_plot(f3o_leg, x = 0.095, y = 0.595-0.5, width = 0.1, height = 0.1) +
  draw_plot(f3q, x = 0.33, y = 0.84-0.5, width = 0.33, height = 0.16) +
  draw_plot(f3r, x = 0.33, y = 0.68-0.5, width = 0.33, height = 0.16) +
  draw_plot(f3t, x = 0.165+.33, y = 0.5-0.5, width = 0.165, height = 0.18) +
  draw_plot(f3s, x = 0+.33, y = 0.5-0.5, width = 0.165, height = 0.18)  +
  draw_plot(f3s_leg, x = 0.095+.33, y = 0.595-0.5, width = 0.1, height = 0.1) +
  draw_plot(f3u, x = 0.66, y = 0.84-0.5, width = 0.33, height = 0.16) +
  draw_plot(f3v, x = 0.66, y = 0.68-0.5, width = 0.33, height = 0.16) +
  draw_plot(f3x, x = 0.165+.66, y = 0.5-0.5, width = 0.165, height = 0.18) +
  draw_plot(f3w, x = 0+.66, y = 0.5-0.5, width = 0.165, height = 0.18) +
  draw_plot(f3w_leg, x = 0.095+.66, y = 0.595-0.5, width = 0.1, height = 0.1)

save_plot(filename = "../manuscript/plots/Figure_3_Data_driven_Covariance_matrices.png", plot = last_plot(), base_height = 10, base_width = 11)
```
